### 2020-02-19 Common/chapters/paper_reproduction.tex
Johan and Julien: section reproducing the paper experiment done

\chapter{Paper reproduction}
\label{ch:paper_reproduction}
This chapter is based on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.\cite{07}, shortly presented in chapter \ref{ch:literature_review}. It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts, which overcome the lack of available data in the medical field (see chapter \ref{ch:transfer_learning}).\\
Song et al. \cite{07} proposed a deep convolutional neural network (DCNN) method to detect prostate cancer based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset. This dataset is one of the biggest available dataset for prostate cancer classification. The two different output classes of the latter are benign lesion (class 0) and malignant lesion (class 1). The dataset contains a total of 204 patients for the training (labels are given) and 208 patients for the challenge (labels are unknown). This paper explains all the steps to follow with the goal of building a computer-aided-diagnosis system for prostate cancer detection. Moreover, it also provides results (AUC) about the algorithm performance on the test set they built (which is not the same as the challenge test set since they split the 204 patients into training,validation/test sets), which can be used as a benchmark to compare with the results of the reproduction of the experiment. However, since the article gives no information about the results their model got on the official PROSTATEx challenge test set (208 patients without labels), the reproduction of the experiment will fill that gap.\\
This chapter is built top-down: first a superficial overview of the entire process is established in order to understand the purpose of the experiment as a whole. Then, each part or theoretical notion of the experiment is described in detail. The structure of the dataset is exposed, all steps of its processing are explained and the techniques used as verification of the proper functioning of the algorithm are presented. After the processing of the data, the training phase is dealt with in depth and all hyperparameters, options and implementation choices are described to ensure the reproductibility of the experiment. This part is followed by the presentation of the raw results the model achieved on the test set (the one built for the experiment), which is itself followed by their analysis in the "discussion" section. Finally, the last section is devoted to the SPIE-AAPM-NCI Prostate MR Classification Challenge and presents the results that the model achieved on the challenge test set using the whole training set for the training. This last part is not present in the original paper and represents contribution of this work to it.

\section{Process overview}
Schematically speaking, the entire experiment process can be represented as shown on figure \ref{fig:paper_reproduction_process}. The first part of the experiment is about the paper reproduction, whereas the second one is devoted to the participation in the SPIE-AAPM-NCI PRO\- STATEx challenge since the authors of the article did not take part in it.\\
The former makes use of the PROSTATEx training set. This dataset is only composed of samples whose clinical significance is provided (labeled data). The first step of the reproduction is the data processing. In fact, from the DICOM files to the numpy arrays that are given to the neural network, many processing steps occur such as the lesion cropping, the normalization the MRI images stacking, the data augmentation, etc (see \ref{sec:prostatex_data_processing} for details). After the processing, the labeled data are split into a training, a validation and a test sets using respectively 80\%, 10\% and 10\% of all available data for each set. The training set is then given in batches as input to the neural network, which updates its weight consequently. At the end of each epoch, the current model is tested on the validation set (that the model has never seen previously) and the metrics are plotted on Tensorboard (see \ref{paper_tensorboard}). If the current model has a bigger AUC and a bigger accuracy than the previous best model, it becomes the best model and it is saved. Furthermore, if the validation AUC is decreasing for a defined number of epoch, the learning rate is reduced to prevent overfitting, which avoids to progressively decrease the performance on the validation set. This technique contributes to maintain a good generalization level for the model and optimized the training with respect to the AUC. When the training phase reaches the defined number of epochs, the model is tested on the test set with their method called "enhanced prediction" and the AUC is computed. This method consists in predicting an output for each augmented image and then computing the average of all predictions for each lesion of each patient. \\
The second part of the experiment relies on the results of the first one. Indeed, the same data processing pipeline is followed and the hyperparameters used to train the neural network are the same but the entire dataset is used as a training set instead of splitting the data into a training/validation/test set. In this way, the entire dataset is given to the neural network, which maximizes the performance of the model on the challenge test set by increasing its generalization ability. During the training, the model is saved at the end of each epoch. At the end of it, multiple models are chosen around the epoch that produced the best result on the validation set during the first part of the experiment. These models are evaluated on the challenge test set according the AUC by sending their predictions for each lesion of each patient. The model with the best result on the challenge is the most efficient one.

@@

\subsection{Results}
Figure \ref{fig:paper_reprodution_results} shows the different metrics registered on the training and validation sets during the training phase. The best model was chosen at epoch 21 with an accuracy of 0.7525 and an AUC of 0.765 (recall: the condition for a model to be chosen as best is that the current AUC and accuracy should be higher than the AUC and accuracy of the previous best model) on the validation set.\\
Figure \ref{fig:paper_reprodution_auc_on_our_test_set}
This best model was also used to take part in the ProstateX challenge. The res XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

@@

\subsection{Discussion}
Thanks to all curves in the "results", a lot of information regarding the model performance can be inferred. In fact, "reviewing learning curves of models during training can be used to diagnose problems with learning, such as an underfit or overfit model, as well as whether the training and validation datasets are suitably representative"\cite{40}.\\
First of all, when observing the loss plot, what is noticeable is that the training and validation losses follow the same behaviour until the epoch 600 approximately. This is a clear sign of good fit of the model on the data. All features learned by the model on the training data are also applicable to the validation set. This situation is the optimal one. After this phase, the model slightly started to overfit. In fact, the training loss continues to linearly decrease whereas the validation loss decreases less quickly until reaching  XXXXXXXXXXXXXXXXXXXXXXXX\\

As the goal of the experiment was to reproduce the results of the paper,

\section{SPIE-AAPM-NCI Prostate MR Classification Challenge}
\label{sec:classification_challenge}
\subsection{Training the neural network with the whole dataset}
\subsection{Results on the challenge test set}

>>>>

\chapter{Paper reproduction}
\label{ch:paper_reproduction}
This chapter is based on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.\cite{07}, shortly presented in chapter \ref{ch:literature_review}. It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts, which overcome the lack of available data in the medical field (see chapter \ref{ch:transfer_learning}).\\
Song et al. \cite{07} proposed a deep convolutional neural network (DCNN) method to detect prostate cancer based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset. This dataset is one of the biggest available dataset for prostate cancer classification. The two different output classes of the latter are benign lesion (class 0) and malignant lesion (class 1). The dataset contains a total of 204 patients for the training (labels are given) and 208 patients for the challenge (labels are unknown). This paper explains all the steps to follow with the goal of building a computer-aided-diagnosis system for prostate cancer detection. Moreover, it also provides results (AUC) about the algorithm performance on the test set they built (which is not the same as the challenge test set since they split the 204 patients into training,validation/test sets), which can be used as a benchmark to compare with the results of the reproduction of the experiment. However, since the article gives no information about the results their model got on the official PROSTATEx challenge test set (208 patients without labels), the reproduction of the experiment will fill that gap.\\
This chapter is built top-down: first a superficial overview of the entire process is established in order to understand the purpose of the experiment as a whole. Then, each part or theoretical notion of the experiment is described in detail. The structure of the dataset is exposed, all steps of its processing are explained and the techniques used as verification of the proper functioning of the algorithm are presented. After the processing of the data, the training phase is dealt with in depth and all hyperparameters, options and implementation choices are described to ensure the reproductibility of the experiment. This part is followed by the presentation of the raw results the model achieved on the test set (the one built for the experiment), which is itself followed by their analysis in the "discussion" section. Finally, the last section is devoted to the SPIE-AAPM-NCI Prostate MR Classification Challenge and presents the results that the model achieved on the challenge test set using the whole training set for the training. This last part is not present in the original paper and represents the contribution of this work to it.

\section{Process overview}
Schematically speaking, the entire experiment process can be represented as shown on figure \ref{fig:paper_reproduction_process}. The first part of the experiment is about the paper reproduction, whereas the second one is devoted to the participation in the SPIE-AAPM-NCI PRO\- STATEx challenge since the authors of the article did not take part in it.\\
The former makes use of the PROSTATEx training set. This dataset is only composed of samples whose clinical significance is provided (labeled data). The first step of the reproduction is the data processing. In fact, from the DICOM files to the numpy arrays that are given to the neural network, many processing steps occur such as the lesion cropping, the normalization the MRI images stacking, the data augmentation, etc (see \ref{sec:prostatex_data_processing} for details). After the processing, the labeled data are split into a training, a validation and a test sets using respectively 80\%, 10\% and 10\% of all available data for each set. The training set is then given in batches as input to the neural network, which updates its weight consequently. At the end of each epoch, the current model is tested on the validation set (that the model has never seen previously) and the metrics are plotted on Tensorboard (see \ref{paper_tensorboard}). If the current model has a bigger AUC and a bigger accuracy than the previous best model, it becomes the best model and it is saved. Furthermore, if the validation AUC is decreasing for a defined number of epoch, the learning rate is reduced to prevent overfitting, which avoids to progressively decrease the performance on the validation set. This technique contributes to maintain a good generalization level for the model and optimizes the training with respect to the AUC. When the training phase reaches the defined number of epochs, the model is tested on the test set with their method called "enhanced prediction" and the AUC is computed. This method consists in predicting an output for each augmented image and then computing the average of all predictions for each lesion of each patient. \\
The second part of the experiment relies on the results of the first one. Indeed, the same data processing pipeline is followed and the hyperparameters used to train the neural network are the same but the entire dataset is used as a training set instead of splitting the data into a training/validation/test set. In this way, the entire dataset is given to the neural network, which maximizes the performance of the model on the challenge test set by increasing its generalization ability. During the training, the model is saved at the end of each epoch. At the end of it, multiple models are chosen around the epoch that produced the best result on the validation set during the first part of the experiment. These models are evaluated on the challenge test set according the AUC by sending their predictions for each lesion of each patient. The model with the best result on the challenge is the most efficient one.

@@

\subsection{Results}
\label{ref:sec:results}
Figure \ref{fig:paper_reprodution_results} shows the different metrics registered on the training and validation sets during the training phase. The best model was chosen at epoch 21 with an accuracy of 0.7525 and an AUC of 0.765 (recall: the condition for a model to be chosen as best is that the current AUC and accuracy should be higher than the AUC and accuracy of the previous best model) on the validation set.\\
This best model was evaluated on the test set built for the experiment and achieved a score of 0.75 in AUC.\\
Finally, the model was also tested on the official SPIE-AAPM-NCI Prostate MR Classification Challenge test set and achieved an AUC of 0.71. This result will be used as a benchmark in comparison with the model trained on the whole dataset in the second part of the experiment.

@@

Thanks to all curves of section \ref{ref:sec:results}, a lot of information regarding the model performance can be inferred. In fact, "reviewing learning curves of models during training can be used to diagnose problems with learning, such as an underfit or overfit model, as well as whether the training and validation datasets are suitably representative"\cite{40}.\\
First of all, when observing the loss plot, what is noticeable is that the training and validation losses have the same behaviour during the 50 epochs. This is a clear sign of good fit of the model on the data. All features learned by the model on the training data are also applicable to the validation set.  This situation is the optimal one. Regarding the slope of the curve, a big change occurs at epoch 17. This moment corresponds to the one where the learning rate is reduced by a factor 100 thanks to the \textit{reducelronplateau}. In fact at this moment, the validation AUC was at its maximum (the training is optimized according to this metric). If this technique had not been used, the training loss would have continued to go down whereas the validation loss would have go up immediately, which would have lead to overfitting. Here thanks to the learning rate decay, this scenario is avoided and the validation and training loss stay stable from this point. The same behaviour is clearly noticeable for all other metrics.\\
Regarding the overall performance of the best model chosen at epoch 21, the latter generalizes well. This is particularly noticeable on the accuracy and AUC plots where the validation metrics are higher or equal to the training ones. The validation accuracy indicates that the model makes the right class decision in approximately 75\% of the time. Furthermore, since the model achieves a validation AUC of 0.76, it shows that it has a good distinction capacity. In fact, the highest is the AUC, the highest is the ability for the model to predict benign lesion as benign and malignant lesion as malignant. Samuel G. et al. \cite{42} claim that a less-experienced radiologist can reach an AUC of 0.81 and an expert an AUC of 0.91. Consequently, this model is almost at the level of a radiologist, which is very promising. It is also important to keep in mind that the model was trained using only 164 patients (80\% of the 204 training patients), which suggests that even better results could be reached with more data. The precision plot, which is used to measure the proportion of malignant lesions correctly classified among all lesions classified as malignant lesions, confirms the idea that the model has a good distinguishable capacity with a score of approximately 0.76 on the validation set. Another extremely important aspect of a computer-aider diagnosis is to avoid at most the number of false negative. In fact, diagnostically, it is far better to classify a benign lesion as malignant than a malignant lesion as benign. This concept is measured by the recall, which quantifies the proportion of malignant lesions correctly classified among all malignant lesion. The model has a recall about 0.74 on the validation set. The two last metrics (precision and recall) are considered together in the f1-score that is equal to 0.76. Finally, the specificity indicates the proportion of correctly classified benign lesions among all benign lesions and is equal to approximately 0.76. Globally, this model is balanced, no metric is exploding at the expense of another and this stability constitutes a serious strength regarding its generalization ability.\\
As the goal of the experiment was to reproduce the experiment of the paper, the comparison between the two models is necessary. The reproducted model achieved an AUC of 0.75 on our test set (see figure \ref{fig:paper_reprodution_auc_on_our_test_set}) using their "enhanced prediction" technique whereas the authors achieved an AUC in the interval [0.876-0.994] with 95\% of confidence on their test set. Consequently, the model presented in this work has a slightly smaller AUC [0.126-0.244] than the one achieved in the paper. Multiple factors can explain this difference. First, the cropping of the lesion was done by a radiologist in their work while it was automatically done using the information in the csv files for this work. Then, the constitution of the test set can make the result vary a lot. In fact, the authors didn't provide any information about how they build their test set. It is therefore possible that the test set was chosen to maximize the performance of the model. In the current work, the test set is built randomly. Consequently, the comparison of the performance is difficult since the test set is not the same in the two works. The only way to seriously compare the two models would be to compare their performance on the official SPIE-AAPM-NCI Prostate MR classification challenge test set. The model of this section (trained with 80\% of all available data) reached an AUC of 0.71. Unfortunately, no information about the performance of the model of the paper is given. Knowing that the best result at the time of the challenge was an AUC of 0.87 on the challenge test set, it is highly surprising not to take part in it when claiming an AUC in the interval [0.876-0.994] with 95\% of confidence... Finally, another source of difference is the number of participants in the experiment and the level of experience they have. Indeed, the paper has seven authors including one bachelors's student and six PhD students with one of them working for Siemens Heathcare and one for hImagingTek Ltd. These big companies are specialized in the medical imaging and have departments dedicated to machine learning. At the opposite, this work was produced independently by two MSc student without previous experience in the medical imaging field.

\section{SPIE-AAPM-NCI Prostate MR Classification Challenge}
\label{sec:classification_challenge}
This section aims at maximizing the result on the SPIE-AAPM-NCI Prostat MR classification Challenge test set by using the entire available data as training set, instead of using 80\% as it was done in the previous section. First, the changes made to reach this goal are detailed, then the raw results are presented and finally the latter are interpreted.
\subsection{Training the neural network with the whole dataset}
\subsection{Results on the challenge test set}
